Concurrent Programming with CXL

### 2020-06-28 book/paper.tex
Latest
Now prints right variables
Added information about HTML output

\begin{figure}
\begin{code}
\begin{verbatim}
#states = 79 diameter = 5
==== Safety violation ====
__init__/() [0,34-49]       50 dict{.count:0, .done:dict{0:False, 1:False}}
incrementer/0 [1-4]          5 dict{.count:0, .done:dict{0:False, 1:False}}
incrementer/1 [1-10]        11 dict{.count:1, .done:dict{0:False, 1:True }}
incrementer/0 [5-10]        11 dict{.count:1, .done:dict{0:True,  1:True }}
main/() [13-17,20-24,26-31] 31 dict{.count:1, .done:dict{0:True,  1:True }}
>>> CXL Assertion failed : 1
Open file://.../cxl.html for more information
\end{verbatim}
\end{code}
\caption{The text output of running CXL on Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/Up1.png}
\caption{The HTML output of running CXL on Figure~\ref{fig:inc}.}
\label{fig:inchtml1}
\end{figure}

@@

@@

Once initialization completes, any processes that were spawned can run.
You can think of CXL as trying every possible interleaving of processes executing
instructions.
Figure~\ref{fig:incoutput} shows the output produced by running CXL on the
\texttt{Up.cxl} program.
It starts by reporting that the assertion on line 9 in column 5 failed and that
\textit{count} has value 1 instead of 0:
\texttt{Assertion failed (’assert’, ’Up.cxl’, 9, 5) 1}.
Next it reports an execution that failed this assertion.  The output has
four columns:
\begin{enumerate}
\item The name tag of the process;
\item The sequence of program counters of the CVM instructions that the process executed;
\item The current program counter of the process;
\item The contents of the shared memory.
\end{enumerate}

If we look at the middle three rows, we see that:
\begin{itemize}
\item Process 0 executed instructions 1 through 4, loading the value of
\textit{count} but stopping just before storing 1 into \textit{count};
\item Process 1 executed instructions 1 through 10, storing 1 into
into \textit{count} and storing \texttt{True} into \textit{done}[1];
\item Process 0 continues execution, storing value 1 into \textit{count}
and storing \texttt{True} into \textit{done}[0].
\end{itemize}

This makes precise the concurrency problem that we encountered.

CXL can report a variety of failure types:
\begin{itemize}
\item \texttt{Safety violation}: this means something went wrong with
at least one of the executions of the program that it tried.  This
can include a failing assertion, divide by zero, using an uninitialized
or non-existent variable, dividing a set by an integer, and so on.
CXL will print at trace of the shortest bad execution that it found.
\item \texttt{Infinite loop}: some process ran into an infinite loop
in which it got back to the same state without having accessed any
shared variables.
\item \texttt{Non-terminating States}: CXL found one or more states
from which there does not exist an execution such that all processes
terminate.  CXL will not only print the non-terminating state with
the shortest trace, but also the list of processes that are running
at that state, along with their name tags and program counters.
\item \texttt{Stopped States}: Similar to non-terminating states,
these are states in which some processes are left that cannot
terminate.  Chapter~\ref{ch:synch} will explain what it means for
a process to be stopped.
\end{itemize}

>>>>

@@

\begin{figure}
\begin{code}
\begin{verbatim}
#states = 79 diameter = 5
==== Safety violation ====
__init__/()   [0,34-49]     50 { count=0, done=[False, False] }
incrementer/0 [1-4]          5 { count=0, done=[False, False] }
incrementer/1 [1-10]        11 { count=1, done=[False, True]  }
incrementer/0 [5-10]        11 { count=1, done=[True,  True]  }
main/() [13-17,20-24,26-31] 31 { count=1, done=[True,  True]  }
>>> CXL Assertion failed : 1
Open file://.../cxl.html for more information
\end{verbatim}
\end{code}
\caption{The text output of running CXL on Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

Once initialization completes, any processes that were spawned can run.
You can think of CXL as trying every possible interleaving of processes executing
instructions.
Figure~\ref{fig:incoutput} shows the output produced by running CXL on the
\texttt{Up.cxl} program.

\begin{figure}
\includegraphics[width=\textwidth]{figures/Up1.png}
\caption{The HTML output of running CXL on Figure~\ref{fig:inc}.
There are four sections.  The top left shows the source code and bytecode
of the program.  The top right shows the steps in an execution.
The middle section shows the values of the variables at a particular step.
The bottom section shows the state of each process.
Each row in the stack trace of a process shows the current program counter,
the method that is being evaluated, and the line of code that is
being evaluated.}
\label{fig:inchtml1}
\end{figure}

CXL can report a variety of failure types:
\begin{itemize}
\item \texttt{Safety violation}: this means something went wrong with
at least one of the executions of the program that it tried.  This
can include a failing assertion, divide by zero, using an uninitialized
or non-existent variable, dividing a set by an integer, and so on.
It also includes certain infinite loops.
CXL will print at trace of the shortest bad execution that it found.
\item \texttt{Non-terminating States}: CXL found one or more states
from which there does not exist an execution such that all processes
terminate.  CXL will not only print the non-terminating state with
the shortest trace, but also the list of processes that are running
at that state, along with their name tags and program counters.
\item \texttt{Stopped States}: Similar to non-terminating states,
these are states in which some processes are left that cannot
terminate.  Chapter~\ref{ch:synch} will explain what it means for
a process to be stopped.
\end{itemize}

In this case, CXL reports a safety violation, in particular it
reports that the assertion on line 9 failed and that
\textit{count} has value 1 instead of 2.
Next it reports an execution that failed this assertion.
The program got to the faulty execution in 5 ``steps.''
The output has
four columns:
\begin{enumerate}
\item The name tag of the process;
\item The sequence of program counters of the CVM instructions that the process executed;
\item The current program counter of the process;
\item The contents of the shared memory.
\end{enumerate}

The first step is initialization.
It sets shared variable \textit{count} to 0 and
shared variable \textit{done} to \texttt{[False, False]}.
If we look at the next three steps, we see that:
\begin{itemize}
\item Process \texttt{implementer/0}
executed instructions 1 through 4, loading the value of
\textit{count} but stopping just before storing 1 into \textit{count};
\item Process \texttt{implementer/1}
executed instructions 1 through 10, storing 1 into
into \textit{count} and storing \texttt{True} into \textit{done}[1];
\item Process \texttt{implementer/0}
continues execution, storing value 1 into \textit{count}
and storing \texttt{True} into \textit{done}[0].
\end{itemize}
Finally, process \texttt{main()} runs and finds the problem.
This makes precise the concurrency issue that we encountered.

CXL also generates an HTML file that allows exploring more details
of the execution interactively.
Open the suggested HTML file and you should see something like
Figure~\ref{fig:inchtml1}.
